---
title: "GSS challange"
author: "Taelynn Condoll"
date: "Summer 2022"
output: html_document
---

```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2018.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

$H_0$:The attitude toward capital punishment is independent of someones race.

$H_A$:The attitude toward capital punishment is dependent of someones race.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

Both race and cappun are categorical variables so analyses technique CAT~CAT, 
with a bar plot, summery is a cross table, inference is chi-squared test.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->



## Now with The Most data (fewer variables gives more data points)

```{r}
dd2 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

According to the bar plot, the race of a person is affected by the way they 
view capital punishment. The graph shows that white people prefer the capital 
punishment and black people strongly disagree with it.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

## Now with More data

```{r}
table2 <- xtabs(~race + cappun, data=mydata)
rowPerc(table2)
colPerc(table2)
```

From the numerical data, white people strongly agree with the death 
sentence, others are slightly in favor of the death sentence and black people 
are lightly against the death sentence. This shows that the attitude 
towards the death sentence vary through race.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

`


## More data version

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
#fisherXY(mydata$race,mydata$cappun)
```

```{r}
fisherXY(mydata$cappun,mydata$race)
```
 
The p value of both chunks is .05. Therefore we can reject the null hypothesis. Also on the fisher exact test the p value is less then .05 as well.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The question was "Is attitude toward capital punishment independent of race?" Through the research provided we can conclude that the race of a person is dependent on their opinion on capital punishment. This was proved with a chi squared test as well as a fisher exact test, as well as a bar plot to depict the level of agreement a white person has on the death sentence then that of other people of color.
